---
title: "HW1"
author: "Xinwei Zhang xz2663"
date: "2/4/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## P3
### (a)

```{r }
set.seed(1)
x = rnorm(100, mean=0, sd=1)
```

### (b)
```{r }
eps = rnorm(100, mean=0, sd=sqrt(0.25))
```

### (c)
```{r }
y = -1 + 0.5*x + eps
```
Length of y is 100. $\beta_0$ is -1, $\beta_1$ is 0.5.

### (d)
```{r }
plot(x,y)
```

There is a linear relationship between x and y with a slope of positive value and variance.

### (e)
```{r}
lm.fit = lm(y~x)
summary(lm.fit)
```

The linear regression fits a model close to the true value of the coefficients as was constructed. The model has a large F-statistic and a near-zero p-value so the null hypothesis can be rejected.

### (f)
```{r}
plot(x,y)
abline(lm.fit, lwd=3, col=2)
abline(-1, 0.5, lwd=3, col=3)
legend(-1, legend = c("model fit", "pop. regression"), col=2:3, lwd=3)
```

### (g)
```{r}
lm.fit_sq = lm(y~x+I(x^2))
summary(lm.fit_sq)
```

There is evidence that this model fit increased given the slight increase in $R^2$ and decrease in $RSE$. Although, the p-value suggests that there isn't a relationship between y and $x^2$.

### (h)
```{r}
set.seed(1)
eps1 = rnorm(100, 0, 0.125)
x1 = rnorm(100)
y1 = -1 + 0.5*x1 + eps1
plot(x1, y1)
lm.fit1 = lm(y1~x1)
summary(lm.fit1)
abline(lm.fit1, lwd=3, col=2)
abline(-1, 0.5, lwd=3, col=3)
legend(-1, legend = c("model fit", "pop. regression"), col=2:3, lwd=3)
```

Observed $RSE$ decreased and $R^2$ increased.

### (i)
```{r}
set.seed(1)
eps2 = rnorm(100, 0, 0.8)
x2 = rnorm(100)
y2 = -1 + 0.5*x2 + eps2
plot(x2,y2)
lm.fit2 = lm(y2~x2)
summary(lm.fit2)
abline(lm.fit2, lwd=3, col=2)
abline(-1, 0.5, lwd=3, col=3)
legend(-1, legend = c("model fit", "pop. regression"), col=2:3, lwd=3)
```

Observed $RSE$ increased and $R^2$ decreased.

### (j)
```{r}
confint(lm.fit)
confint(lm.fit1)
confint(lm.fit2)
```

All intervals centered around 0.5. The interval of fit1 (less noisy) is narrower than that of fit (original). The interval of fit2 (noisier) is wider than that of fit (original).

## P4
```{r}
library(ISLR)
Advertising <- read.csv("http://www-bcf.usc.edu/~gareth/ISL/Advertising.csv",
                        header=T,na.string=',')
#attach(Advertising)
model_newspaper = lm(sales~newspaper, data=Advertising)
model_TV = lm(sales~TV, data=Advertising)
model_radio = lm(sales~radio, data=Advertising)
confint(model_newspaper, , 0.92)
confint(model_TV, , 0.92)
confint(model_radio, , 0.92)
plot(Advertising$newspaper,Advertising$sales, xlab = "newspaper", ylab = "sales", 
     main = "sales versus newspaper")
abline(model_newspaper, lwd=3, col=2)
confint_newspaper = predict(model_newspaper, newspaper = data.frame(Advertising$newspaper), 
                            interval = "confidence", level=0.92)
lines(Advertising$newspaper, confint_newspaper[,2], col="purple")
lines(Advertising$newspaper, confint_newspaper[,3], col="purple")
plot(Advertising$TV,Advertising$sales, xlab = "TV", ylab = "sales", 
     main = "sales versus TV")
abline(model_TV, lwd=3, col=2)
confint_TV = predict(model_TV, TV = data.frame(Advertising$TV), 
                     interval = "confidence", level=0.92)
lines(Advertising$TV, confint_TV[,2], col="purple")
lines(Advertising$TV, confint_TV[,3], col="purple")
plot(Advertising$radio,Advertising$sales, xlab = "radio", ylab = "sales", 
     main = "sales versus radio")
abline(model_radio, lwd=3, col=2)
confint_radio = predict(model_radio, radio = data.frame(Advertising$radio), 
                        interval = "confidence", level=0.92)
lines(Advertising$radio, confint_radio[,2], col="purple")
lines(Advertising$radio, confint_radio[,3], col="purple")
```

### P5
### (a)
```{r}
pairs(Auto)
```

### (b)
```{r}
cor(subset(Auto, select=-name))
```

### (c)
```{r}
lm.fit1 = lm(mpg~.-name, data=Auto)
summary(lm.fit1)
```

### (i)
There is a relationship between the predictors and the response by testing the null hypothesis of whether all the regression coeffcients are zero. F statistic is large with small p value, which is againest the null hypothesis.

### (ii)
Displacement, weight, year and origin have significant relationship to the response.

### (iii)
The coefficient for year, 0.750773, suggests that for every year, mpg increases by 0.750773. Cars become more fuel efficient every year.

### (d)
```{r}
lm.fit2 = lm(mpg~log(weight)+sqrt(horsepower)+I(acceleration^2), data=Auto)
summary(lm.fit2)
```

From p values, log(weight) and sqrt(horsepower) have significant relationship to mpg. 




